---
title: 'A01: GGPLOT2'
author: Caleb Coulson
date: '2021-05-18'
slug: a01-ggplot2
categories:
  - ggplot2
tags: []
---
## This post contains two data visualizations, and are as follows

* one Numerical visualization
* one Categorical data visualization

## The following data was used to create the data visualizations that follow


```{r message= FALSE, echo=FALSE}
library(dplyr)
mtcars 
```

## The first Data visualization concerns ata was extracted from the 1974 Motor Trend US magazine and relates the weight of vehicles to the miles per gallon. 



```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y= mpg),colour(cyl)) +
  geom_point() + labs(title = "Miles per Gallon vs. weight of vehicles from 1974", x = "weight(1000lb)", y = "Miles per Gallon")
```

### From the previous plot it can be seen that as the weight of the vehicle increases, the miles per galon decreases, as is to be expected.

## The second data visualization is categorical and defines the number of cylinders in the engines of the vehicles of 1974

```{r echo=FALSE}
library(ggplot2)
ggplot(mtcars, aes(x = cyl,fill = vs)) +
  geom_bar()+ labs(title = "Number of cyclinders in engines of cars from 1974", x = "Number of Cylinders", y = "Count")
```

### From the second plot it can be seen that the most popular category of "engine cylinders" is an 8 cylinder. 
