---
title: "A03 Assignment - Caleb Coulson"
author: "Caleb Coulson"
date: '2021-06-03'
slug: a03-assignment-caleb-coulson
categories: []
tags: []
---

## inporting data and loading librarys
```{r}
library(tidyverse)
library(readxl)
lego_sales <- read_csv("lego_sales.csv")
print(lego_sales)
```

## most common first names 
```{r}
lego_sales %>%
  count(first_name) %>%
  arrange(desc(n))
```
### From this it is apparant that "Jackson" is the most common first name with a count of 13

## most common theme
```{r}
lego_sales %>%
  count(theme) %>%
  arrange(desc(n))
```
### From this it is apparant that "star wars" is the most common theme with a count of 75


## most common sub-theme for star wars
```{r}
lego_sales %>%
  filter(theme == 'Star Wars') %>%
  count(subtheme) %>%
  arrange(desc(n))
```
### From this it is apparant that "The Force Awakens" is the most common subtheme within the star wars theme with a count of 15

## creating new variable "age_group"
```{r}
updated_lego_sales <- lego_sales %>%
  mutate(age_group = case_when(
    age <= 18 ~ '18 and under',
    between(age, 19, 25) ~ '19 - 25',
    between(age, 26, 35) ~ '26 - 35',
    between(age, 36, 50) ~ '36 - 50',
    TRUE ~ '51 and over'))
```

## answering Question 5 probabilities
```{r}
## part a
updated_lego_sales %>%
  count(age_group) %>%
  mutate(prob_a=(n/sum(n)))

## part b
updated_lego_sales %>%
  count(age_group, theme) %>%
 mutate(prob_b=(n/sum(n)))
 
## part c
updated_lego_sales %>%
  count(age_group, theme) %>%
filter(theme=='Duplo') %>%
  mutate(prob_c=(n/sum(n)))

```
### From this it can be seen that the probability..
### a randomly selected customer is in the 19 - 25 age group = 0.2080645
### a randomly selected customer is in the 19 - 25 age group and purchased a Duplo theme set = 0.014516129
###  a randomly selected customeris in the 19 - 25 age group given they purchased a Duplo theme set = 0.25714286	

## part 6 (Which age group has purchased the largest number of lego sets? How many did they purchase)
```{r}
updated_lego_sales %>%
  count(age_group)
```
### from this it can be seen that the age group that has purchased the most lego sets is the 36-50 age group, and they purchaced 216 sets

## part 7 (Which age group has spent the most money on legos? How much did they spend?)
```{r}
updated_lego_sales %>%
  count(age_group,us_price) %>%
  mutate(total_cost = us_price*n) %>%
   group_by(age_group) %>%
   mutate(cost_sum= sum(total_cost))
```
### from this it can be seen that the 36-50 age group spent the most money on legos and spent a total of $6641.84


## Part 8 - The question i want to see answered is how the ammounnt of pieces affects the price of the set, and Im also going to color it by age_group to see if there is a trend in what price lego sets the age groups buy.
```{r}
ggplot(updated_lego_sales, aes(x=pieces, y=us_price, color=age_group)) +
  geom_point() 
```

## part 9 - i have changed both the x and y axis to log scales to spread out the data more, in addition to this I have changed the theme to a dark version because I like this more for aestetics.
```{r}
ggplot(updated_lego_sales, aes(x=pieces, y=us_price, color=age_group)) +
  geom_point() +
  scale_x_log10()+
  scale_y_log10() + 
  theme_dark()
```

