---
title: 'Assignment A02: TIDYVERSE'
author: CalebCoulson
date: '2021-05-27'
slug: assignment-a02-tidyverse
categories: []
tags: []
---
```{r}
library(dplyr)
library(ggplot2)
library(gapminder)
library(tidyr)
```


## Inporting data from 2017 and 2020 PGA tour season
```{r}
library(tidyverse)
golf2017 <- read_csv("PGATOUR_data2.csv")
head(golf2017)
names(golf2017)
library(tidyverse)
golf2020 <- read_csv("pga_tour_stats_2020.csv")
head(golf2020)
names(golf2020)
```

## Determining The % of top 10 finishes for players who played 20 or more events
```{r}
golf2017UP <- golf2017 %>%
  group_by(NUMBER_OF_TOP_Tens) %>%
 filter(NUMBER_OF_TOP_Tens >= 3) %>%
  mutate(TOP10_percent = NUMBER_OF_TOP_Tens/EVENTS_PLAYED)
```

## Plotting the relationship of top 10 finishes to average driving distance
```{r}
ggplot(golf2017UP, aes(x=AVG_Driving_DISTANCE, y=TOP10_percent)) +
  geom_point()

```

## Adding Player column to match maning convention of 2017 stats
```{r}
golf2020UP <- golf2020 %>%
  mutate(Player = `PLAYER NAME`)
```


## Combining all data
```{r}
golf2017UP %>%
  full_join(golf2020UP, by = 'Player')
```

## Adding 2017 and 2020 PGA tour stats using inner join, to show who was on tour both years, and comparing their club head speeds to see who added most speed to driver.
```{r}
golf2017UP %>%
  inner_join(golf2020UP, by = 'Player') %>%
mutate(speed_change = (TEE_AVG_CLUB_HEAD_SPEED-AVG_CLUB_HEAD_SPEED)) %>%
  select(Player,speed_change) %>% 
  arrange(desc(speed_change))
```

## Comparing top 10 finishers to their odds of getting ball up and down around green
```{r}
golf2020 %>%
  group_by(FINISHES_TOP10) %>%
  summarise(UP_DOWN_PERC=mean(GREEN_PCT_SCRAMBLE_FRINGE)) %>%
  drop_na(FINISHES_TOP10) %>%
  arrange(desc(FINISHES_TOP10))
```

## comparing age of player to events played in 2017
```{r}
golf2017 %>%
pivot_longer(
  cols = AGE,
  names_to = "AGE OF PLAYER",
  values_to = "years"
) %>%
  select(Player, 'AGE OF PLAYER', years, EVENTS_PLAYED) %>%
  drop_na(EVENTS_PLAYED) %>%
arrange(desc(years))

```
## CREATING SUMMARY TABLE OF AVERAGE ONE, TWO, AND THREE PUTTS FOR 18 HOLES FOR EACH PLAYER IN THE PGA IN 2020
```{r}
golf2020 %>%
pivot_longer(
  cols = c(PUTTING_AVG_ONE_PUTTS, PUTTING_AVG_TWO_PUTTS, PUTTING_AVG_THREE_PUTTS),
  names_to = "putting averages",
  values_to = "value"
  ) %>%
  select('PLAYER NAME','putting averages', value)
```

#### while I was working on this assignment I selected the PGA tour stats because I am a big fan of golf and wanted to see if I could get some good data corrilaton. However I was suprised to find out hat no single stat was a good predictor of ho would win or come top 10 the most in a single season. I struggled learning how to properly use the tidying (pivot) functions but figured it out eventually! I follow the PGA closely and I know that bryson DEechambeau has added a lot of distance to his driver so i used this as a check to ensure my code was right, when i saw he was at the top of the list i knew my code was correct. I was suprised to see that the age of the golfer did not correlate much to the number of events played, was anticipating the older players to play much less however this was not the case. Overall i enjoyed this assignemnt but it was a lot harder than the last one.

